---
title: "EEMB 146 Final Report: Example with Iris Data"
author: "Tatum Katz"
date: "5/13/2020"
output: pdf_document
---

**This is a merely a template or the "bare bones" of what your final report should look like. I reccomend copying and pasting this into a .Rmd in your .Rproj. Once pasted in your .Rmd, I reccomend filling in the sections with your writing and code. Pay attention to the document format and figure labels within {r}.**


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE) # you must have this set to TRUE

```


```{r data and libraries, include=FALSE}
# upload necessary libraries
library(readr)
library(tidyverse)
library(ggplot2)
library(psych)
library(car)
library(multcomp)
library(bookdown)

# include other libraries you may need

# upload your dataset(s)
dat <- iris #if you needed to clean or transform the data, it should be done up here so when you make the plots below they work.

```

## Abstract

Here, I would write ~4 sentences about why this research is relevant, what I did, and my findings.

## Introduction

Here, I would give a general overview of my dataset; the variables and how it was collected. I would also talk about my research question, and why it is interesting and important.

## Exploratory Data Analysis

Here, I would mostly state the findings that the petal variables are nonnormal, and references the figures I made. Check out the R code to see how I got these figures to show up where I wanted them to! (Fig. 1). Then, I would reference Figure 2, which is a nice example of a use of the function par. Hey, did you notice it automatically numbers my figures in the order they appear?

```{r warnings=FALSE, fig.show="hold", echo=FALSE, output=FALSE, fig.cap="i would write a VERY informative figure caption here that explained EVERY THING about this plot!"}

#more numeric var investigation
pairs.panels(dat[,-5]) #include as a fig

```


```{r warnings=FALSE, fig.show="hold", echo=FALSE, output=FALSE, fig.cap="i would write a VERY informative figure caption here that explained EVERY THING about this plot!"}

par(mfrow=c(2,2)) #include as a fig
boxplot(Sepal.Length ~ Species, data=dat, ylab="Sepal Length", xlab="Species")
boxplot(Sepal.Width ~ Species, data=dat, ylab="Sepal Width", xlab="Species")
boxplot(Petal.Length ~ Species, data=dat, ylab="Petal Length", xlab="Species")
boxplot(Petal.Width ~ Species, data=dat, ylab="Petal Width", xlab="Species")
par(mfrow=c(1,1))

```

## Statistical Methods

Here, I would explain generally what I did, and then make subsections!

#### Does petal length vary by species?

I would CLEARLY state my hypotheses, and the test I ran. I would describe what ANOVA is, what its assumptions are and how I checked them, You could even get mathy here if you wanted to with some $latex$, but it isn't required.

#### What variables are good predictors of petal length?

Same as above, but you could get super fancy with some $$latex on its own line using double dollar signs $$.

## Results

Now, I would generally explain my results, and go for those subsections again bc i love me some subsections.

#### Does petal length vary by species?

I would report all my statistical results like this: I found a significant difference across species' mean petal lengths (one-way ANOVA; F-value = 1180, p-value < 0.05). And then reference Figure 3!

```{r warnings=FALSE, fig.show="hold", echo=FALSE, output=FALSE, fig.cap="i literally did all my analyses in the appendix and then copied and pasted my favorite plots up here so they show up in the main body!"}
label.df <- data.frame(Species = levels(dat$Species), #make significance stars
                       Petal.Width = c(.7, 1.9, 2.6)) #these are the y-axis locs of the stars

ggplot(dat, aes(x=Species, y=Petal.Width, fill=Species, color=Species))+ #base plot
  geom_boxplot(show.legend=F)+ #do a boxplot, but hide legend cause its just colors
  scale_fill_viridis_d()+ #set the fill colors
  scale_color_manual(values=c("orchid", "lightseagreen", "orange"))+ #set line colors
  ylab("Petal Width")+ 
  geom_text(data = label.df, label = "***", size=10, show.legend=F)+ #add those ***
  theme_bw()+ #cute theme
  theme(text = element_text(size=18)) #cause my eyes is tired and old

```

#### What variables are good predictors of petal length?

Rinse, and repeat (Figure 4).

```{r warnings=FALSE, fig.show="hold", echo=FALSE, output=FALSE, fig.cap="i literally did all my analyses in the appendix and then copied and pasted my favorite plots up here so they show up in the main body!"}

ggplot(dat, aes(x=Sepal.Length, y=Petal.Length, color=Species, fill=Species))+
  geom_jitter(size=3, alpha=0.6, shape=21)+
  geom_smooth(method="lm", alpha=0.6, size=0.5)+
  scale_color_manual(values=c("orchid", "lightseagreen", "orange"))+
  scale_fill_viridis_d()+ #set the fill colors
  ylab("Petal Width")+ 
  xlab("Sepal Length")+
  theme_bw()+ #cute theme
  theme(text = element_text(size=18)) #cause my eyes is tired and old

```

## Discussion

Here, I would wrap up everything and explain the biological signficiance. What didn't I do? What were the limitations? What would I do next?

## References

Site each paper and package here. Try using the citation function to get your cites for each package! 

## Appendix

### Exploratory data analysis

```{r}

#general data checking
head(dat)
levels(dat$Species)
summary(dat)

#check numeric vars for distributions and normality
par(mfrow=c(2,2))
hist(dat$Sepal.Length, main="", xlab="Sepal Length")
hist(dat$Sepal.Width, main="", xlab="Sepal Width")
hist(dat$Petal.Length, main="", xlab="Petal Length")
hist(dat$Petal.Width, main="", xlab="Petal Width")
par(mfrow=c(1,1))

shapiro.test(dat$Sepal.Length) #not normal
shapiro.test(dat$Sepal.Width) #normal
shapiro.test(dat$Petal.Length) #not normal
shapiro.test(dat$Petal.Width) #not normal

par(mfrow=c(2,2))
qqPlot(dat$Sepal.Length) #looks normal
qqPlot(dat$Sepal.Width) #looks normal
qqPlot(dat$Petal.Length) #looks not normal
qqPlot(dat$Petal.Width) #looks not normal
par(mfrow=c(1,1))

#more numeric var investigation
pairs.panels(dat[,-5]) #include as a fig

#numeric vars vs. species
par(mfrow=c(2,2)) #include as a fig
boxplot(Sepal.Length ~ Species, data=dat, xlab="Sepal Length")
boxplot(Sepal.Width ~ Species, data=dat, xlab="Sepal Width")
boxplot(Petal.Length ~ Species, data=dat, xlab="Petal Length")
boxplot(Petal.Width ~ Species, data=dat, xlab="Petal Width")
par(mfrow=c(1,1))

#based on all this, here are my findings:

#petal vars not normal, may need to be transformed
#species seems like a good predictor of each numeric var, perhaps an ANOVA, or multivariate linear model
#sepal length and petal length look correlated
#sepal length and petal width look correlated
#petal length and petal width look correlated

#so, i think i will run an anova on petal length ~ species for my "comparing means" test
#and run a linear model to see what variables are most predictive of petal length for my "predictive" test

```


### Statistical methods
#### ANOVA
```{r}

#goal: one way anova of petal length by species

#step 1: check assumptions of anova
#indep data? probs
#normality
#equal variance

fit1 <- lm(Petal.Length ~ Species, data=dat)
par(mfrow=c(2,2))
plot(fit1) #oof looks a little cone shaped
par(mfrow=c(1,1))
res1 <- fit1$residuals
shapiro.test(res1) #just barely non-normal
qqPlot(res1) #actually looks normal
hist(res1) #hmm, not obviously skewed

#so the variances aren't great but the data are normal enough so maybe i can ignore that violation!

#step 2: run the anova

irisaov <- aov(Petal.Length ~ Species, data=dat)
summary(irisaov) #species highly significant, as expected from EDA. only added digits arg bc otherwise pvals were rounded to 0

#step 3: post-hoc analyses

TukeyHSD(irisaov) #whoa p-values should never be 0. this is a rounding error. lets extract with more precision
thsd <- TukeyHSD(irisaov)
print(thsd, digits=15) #wow so super significant, they are all sig different from each other

#step 4: make a nice ggplot figure to summarize these results, include as a fig

label.df <- data.frame(Species = levels(dat$Species), #make significance stars
                       Petal.Width = c(.7, 1.9, 2.6)) #these are the y-axis locs of the stars

ggplot(dat, aes(x=Species, y=Petal.Width, fill=Species, color=Species))+ #base plot
  geom_boxplot(show.legend=F)+ #do a boxplot, but hide legend cause its just colors
  scale_fill_viridis_d()+ #set the fill colors
  scale_color_manual(values=c("orchid", "lightseagreen", "orange"))+ #set line colors
  ylab("Petal Width")+ 
  geom_text(data = label.df, label = "***", size=10, show.legend=F)+ #add those ***
  theme_bw()+ #cute theme
  theme(text = element_text(size=18)) #cause my eyes is tired and old

```

#### Linear regression
```{r}

#goal: predict petal length by all, see what is important

#step 1: check assumptions of linear regression
#indep? i hope so
#normal residuals?
#equal variances?

fit2 <- lm(Petal.Length ~ Sepal.Length + Species, data=dat) 
par(mfrow=c(2,2))
plot(fit2) #looks homoskedastic to me
par(mfrow=c(1,1))
res2 <- fit2$residuals
qqPlot(res2) #she is normal
shapiro.test(res2) #yep its normo

#step 2: run the lm

summary(fit2) #EVERYTHING! is a significant predictor, which isn't shocking based on EDA

#step 3: make a pretty graph to summarize findings

ggplot(dat, aes(x=Sepal.Length, y=Petal.Length, color=Species, fill=Species))+
  geom_jitter(size=3, alpha=0.6, shape=21)+
  geom_smooth(method="lm", alpha=0.6, size=0.5)+
  scale_color_manual(values=c("orchid", "lightseagreen", "orange"))+
  scale_fill_viridis_d()+ #set the fill colors
  ylab("Petal Width")+ 
  xlab("Sepal Length")+
  theme_bw()+ #cute theme
  theme(text = element_text(size=18)) #cause my eyes is tired and old

```








