---
title: "FinalProject_DatasetGenerator"
author: "sbsambado"
date: "4/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(xlsx)
library(ggplot2)
library(dplyr)
```

Final Project datasets

Datasets **should include at a minimum**
- 1 numeric normally distributed variable (y outcome)
- 1 numeric variable (x input)
- 2 categorical variables (x inputs)

Datasets **should not include**
- time series focused data
- aggregated health data


#### Open Traits Network
This is a solid database that has ecological datasets students are used to seeing in this class.

For each solid database (i.e. AmphiBIO), I would make 5 unique datasets from by randomly selecting 300 observations for each dataset. The goal of choosing 300 observations is to ensure normality should be met and avoid too much incomplete data. 

If student is struggling to meet 1) normality and has 2) a lot of missing data with a particular dataset, change the random number generator to 500.
> sample_n(DATASET, 500)

Dataset: AmphiBIO
```{r}
# https://opentraits.org/datasets/amphi-bio
amphiBIO <- read_csv("AmphiBIO_v1.csv")
dim(amphiBIO)

dataset_1 <- sample_n(amphiBIO, 300)
dim(dataset_1)
write.xlsx(dataset_1, file = "dataset_1.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_15 <- sample_n(amphiBIO, 300)
write.xlsx(dataset_15, file = "dataset_15.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_30 <- sample_n(amphiBIO, 300)
write.xlsx(dataset_30, file = "dataset_30.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_45 <- sample_n(amphiBIO, 300)
write.xlsx(dataset_45, file = "dataset_45.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_60 <- sample_n(amphiBIO, 300)
write.xlsx(dataset_60, file = "dataset_60.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_75 <- sample_n(amphiBIO, 300)
write.xlsx(dataset_75, file = "dataset_75.xlsx",
           col.names = TRUE, row.names = TRUE)

```

Dataset: Arthropod Species Traits
```{r}
# https://opentraits.org/datasets/arthropod-species-traits

arthropod <- read_csv("ArthropodSpeciesTraits.csv")
dim(arthropod)

dataset_2 <- sample_n(arthropod, 300)
dim(dataset_2)
write.xlsx(dataset_2, file = "dataset_2.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_16 <- sample_n(arthropod, 300)
write.xlsx(dataset_16, file = "dataset_16.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_31 <- sample_n(arthropod, 300)
write.xlsx(dataset_31, file = "dataset_31.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_46 <- sample_n(arthropod, 300)
write.xlsx(dataset_46, file = "dataset_46.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_61 <- sample_n(arthropod, 300)
write.xlsx(dataset_61, file = "dataset_61.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_76 <- sample_n(arthropod, 300)
write.xlsx(dataset_76, file = "dataset_76.xlsx",
           col.names = TRUE, row.names = TRUE)

```

Dataset: Australian birds
```{r}
# https://opentraits.org/datasets/australian-birds

australian <- read_csv("Australian_Bird_Data_Version_1.csv")
dim(australian)

dataset_3 <- sample_n(australian, 300)
dim(dataset_2)
write.xlsx(dataset_3, file = "dataset_3.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_17 <- sample_n(australian, 300)
write.xlsx(dataset_17, file = "dataset_17.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_32 <- sample_n(australian, 300)
write.xlsx(dataset_32, file = "dataset_32.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_47 <- sample_n(australian, 300)
write.xlsx(dataset_47, file = "dataset_47.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_62 <- sample_n(australian, 300)
write.xlsx(dataset_62, file = "dataset_62.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_77 <- sample_n(australian, 300)
write.xlsx(dataset_77, file = "dataset_77.xlsx",
           col.names = TRUE, row.names = TRUE)


```

Dataset: Avian hand-wing index
```{r}
# https://opentraits.org/datasets/avianhwi

handwing <- read_csv("Australian_Bird_Data_Version_1.csv")
dim(handwing)

dataset_4 <- sample_n(handwing, 300)
dim(dataset_4)
write.xlsx(dataset_4, file = "dataset_4.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_18 <- sample_n(handwing, 300)
write.xlsx(dataset_18, file = "dataset_18.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_33 <- sample_n(handwing, 300)
write.xlsx(dataset_33, file = "dataset_33.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_48 <- sample_n(handwing, 300)
write.xlsx(dataset_48, file = "dataset_48.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_63 <- sample_n(handwing, 300)
write.xlsx(dataset_63, file = "dataset_63.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_78 <- sample_n(handwing, 300)
write.xlsx(dataset_78, file = "dataset_78.xlsx",
           col.names = TRUE, row.names = TRUE)
```

Dataset: Beukhof_2019 - Marine Fish
```{r}
# https://opentraits.org/datasets/beukhof-2019

marinefish <- read_csv("Beukhof_Fish.csv")
dim(marinefish)

dataset_5 <- sample_n(marinefish, 300)
dim(dataset_5)
write.xlsx(dataset_5, file = "dataset_5.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_19 <- sample_n(marinefish, 300)
write.xlsx(dataset_19, file = "dataset_19.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_34 <- sample_n(marinefish, 300)
write.xlsx(dataset_34, file = "dataset_34.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_49 <- sample_n(marinefish, 300)
write.xlsx(dataset_49, file = "dataset_49.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_64 <- sample_n(marinefish, 300)
write.xlsx(dataset_64, file = "dataset_64.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_79 <- sample_n(marinefish, 300)
write.xlsx(dataset_79, file = "dataset_79.xlsx",
           col.names = TRUE, row.names = TRUE)

```

Dataset: Bland_2017 - Cray Fish
```{r}
# https://opentraits.org/datasets/bland-2017

crayfish <- read_csv("Bland_2017_Crayfish_Species_Dataset.csv")
dim(crayfish)

dataset_6 <- sample_n(crayfish, 300)
dim(dataset_6)
write.xlsx(dataset_6, file = "dataset_6.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_20 <- sample_n(crayfish, 300)
write.xlsx(dataset_20, file = "dataset_20.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_35 <- sample_n(crayfish, 300)
write.xlsx(dataset_35, file = "dataset_35.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_50 <- sample_n(crayfish, 300)
write.xlsx(dataset_50, file = "dataset_50.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_65 <- sample_n(crayfish, 300)
write.xlsx(dataset_65, file = "dataset_65.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_80 <- sample_n(crayfish, 300)
write.xlsx(dataset_80, file = "dataset_80.xlsx",
           col.names = TRUE, row.names = TRUE)
```


Dataset: EGrowth Worms
```{r}
# https://opentraits.org/datasets/egrowth


eGrowth <- read_csv("eGrowth.csv")
dim(eGrowth)

dataset_7 <- sample_n(eGrowth, 300)
dim(dataset_7)
write.xlsx(dataset_7, file = "dataset_7.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_21 <- sample_n(eGrowth, 300)
write.xlsx(dataset_21, file = "dataset_21.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_36 <- sample_n(eGrowth, 300)
write.xlsx(dataset_36, file = "dataset_36.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_51 <- sample_n(eGrowth, 300)
write.xlsx(dataset_51, file = "dataset_51.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_66 <- sample_n(eGrowth, 300)
write.xlsx(dataset_66, file = "dataset_66.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_81 <- sample_n(eGrowth, 300)
write.xlsx(dataset_81, file = "dataset_81.xlsx",
           col.names = TRUE, row.names = TRUE)


```


Dataset: Bacterial and archaeal phenotypic trait database
```{r}
# https://opentraits.org/datasets/madin-2020


bacteria <- read_csv("Bacteria_condensed_traits_GTDB.csv")
dim(bacteria)

dataset_8 <- sample_n(bacteria, 300)
dim(dataset_8)
write.xlsx(dataset_8, file = "dataset_8.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_22 <- sample_n(bacteria, 300)
write.xlsx(dataset_22, file = "dataset_22.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_37 <- sample_n(bacteria, 300)
write.xlsx(dataset_37, file = "dataset_37.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_52 <- sample_n(bacteria, 300)
write.xlsx(dataset_52, file = "dataset_52.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_67 <- sample_n(bacteria, 300)
write.xlsx(dataset_67, file = "dataset_67.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_82 <- sample_n(bacteria, 300)
write.xlsx(dataset_82, file = "dataset_82.xlsx",
           col.names = TRUE, row.names = TRUE)


```


Dataset: Senescence / Aging and longevity of animals

```{r}
# https://opentraits.org/datasets/senescence


senescence <- read_csv("Senescence_AgingAnimals.csv")
dim(senescence)

dataset_9 <- sample_n(senescence, 300)
dim(dataset_9)
write.xlsx(dataset_9, file = "dataset_9.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_23 <- sample_n(senescence, 300)
write.xlsx(dataset_23, file = "dataset_23.xlsx",
           col.names = TRUE, row.names = TRUE)
          
dataset_38 <- sample_n(senescence, 300)
write.xlsx(dataset_38, file = "dataset_38.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_53 <- sample_n(senescence, 300)
write.xlsx(dataset_53, file = "dataset_53.xlsx",
           col.names = TRUE, row.names = TRUE)


dataset_68 <- sample_n(senescence, 300)
write.xlsx(dataset_68, file = "dataset_68.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_83 <- sample_n(senescence, 300)
write.xlsx(dataset_83, file = "dataset_83.xlsx",
           col.names = TRUE, row.names = TRUE)


```


#### Tidy Tuesday Data

Tried to get some fun environmental data here but it takes a while to find appropriate datasets that meet all project requirements

```{r}
install.packages("tidytuesdayR")
library(tidytuesdayR)
```

Transit Costs Project
```{r}
transit_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
write.xlsx(transit_cost, file = "transit_cost.xlsx",
           col.names = TRUE, row.names = TRUE)

dim(transit_cost)


dataset_10 <- sample_n(transit_cost, 300)
dim(dataset_10)
write.xlsx(dataset_10, file = "dataset_10.xlsx",
           col.names = TRUE, row.names = TRUE)

dataset_69 <- sample_n(transit_cost, 300)
write.xlsx(dataset_69, file = "dataset_69.xlsx",
           col.names = TRUE, row.names = TRUE)
```

Deforestation
```{r}
brazil_loss <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/brazil_loss.csv')

```

