---
title: "Lab2_answers"
author: "sbsambado"
date: "1/16/2021"
output:
  html_document: default
  pdf_document: default
---

#Answers for Lab 2

**Question 1** (6 points: 2 points code is correct, 1 point for correct response)

+ A. The ground squirrel has the largest brain size:body ratio of 40 g/kg.

+ B. The African elephant has the smallest brain size:body ratio of 0.858431 g/kg.

+ C. The mean brain size:body ratio is 9.516999 g/kg.  This is less than the human ratio of 21.29032.

+ D. Anything coherent that explains their findings about the data not supporting scientist hypothesis in complete sentencees.


**Question 2** (2 points: 1 point for code, 1 point for response)

+ There are 16 species of desert birds that are considered rare. 

**Question 3** (2 points: 1 point for code, 1 point for response)

+ 27% of endangered species are vertebrates.


*Notes*

+ ALL HOMEWORK SHOULD BE SUBMITTED AS A KNITTED HOMEWORK FILE (.PDF or .HTML)

  + Reduce 20% off HW not in right format (-2 points)

+ Take the time to nail in homework format since this is first graded assignment
  + -.25 point if code is not annotated
  + -.25 point if answers are not in complete sentences
  + -.25 point if code does not seem adequate to answer question




Some potential code to answer questions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(knitr)
library(ggplot2)
```

**Question 1.**
```{r}
# load in required data
mammal <- read_csv("mammals.csv")

# let's look at file
#View(mammal)

names(mammal) <- c("name", "body", "brain")

# A. Which mammal has the largest brain size:body ratio?

# make new variable column
mammal$ratio <- mammal$brain / mammal$body

# / divide 
# * multiply 
# exp is exponential 


# find max ratio
# dataframe[which.function(df$variable),]

mammal[which.max(mammal$ratio),] # don't forget the comma post ()!!!

# B. Which mammal has the smallest ratio?
mammal[which.min(mammal$ratio),] # don't forget the comma post ()!!!

# C. What is the mean brain size:body ratio? Are humans above or below the mean?
mean(mammal$ratio)

# find human ratio

# Option 1 (prefered)
mammal[mammal$name == "Human",] # make sure the spelling is the exact same!

# Option 2
mammal[mammal$name == "Human", "ratio"] # if you have a lot of variable columns

# Option 3 (less prefered)
# make new dataframe for humans and analyze that
human <- mammal[which(mammal$name == "Human"),]
mean(human$ratio)
```

**Question 2**
```{r}
# load data

birds <- read_csv("desert_birds.csv")
#View(birds)

# A. How many desert bird species are rare (<= 10 counts)

# make subset

rare <- subset(birds, Count <= 10)
#find how many rows there are in subset

# Option 1. Look at length of vector
length(rare$Count)
dim(rare)
nrow(rare)

# Option 2. Look at dataset summary
print(rare)
summary(rare)

# This is different than finding the total sum of birds within these rare species
sum(rare$Count)
```

**Question 3**
```{r}

# load data
endangered <- read_csv("endangered.csv")

#View(endangered)

# A. What proportion of endangered species are vertebrates?

# Option 1

Verts <- subset(endangered, vertebrate == "yes")
sum(Verts$no.species) / sum(endangered$no.species) # 0.26

# / divide 
# * multiple
# exp is exponential 


# Option 2

verts <- sum(endangered$no.species[endangered$vertebrate == "yes"])
prop <- verts/sum(endangered$no.species)
print(prop) # 0.27

# Option 3
# make new dataframe

vertebrate <- subset(endangered, vertebrate == "yes")

# find sum of vertebrates
vertebrate_count <- sum(vertebrate$no.species)
print(vertebrate_count) # 399

# find sum of all endangered species (verts & inverts)
endangered_count <- sum(endangered$no.species)
print(endangered_count) # 1263

# find proportion of vertebrates
proportion <- vertebrate_count/endangered_count
print(proportion) # 0.27




```
