---
title: "Lab2_RmarkdownDataAnalysis"
author: "sbsambado"
date: "1/14/2021"
output: html_document
---
Lab 2: Introduction to R Markdown and Data Analysis

# 3 common syntax errors
# spelling 
install.packages('readr'), correct
instal.package('redr') , not correct

# capitilization
install.packages('redr'), correct
install.packages('redr'), not correct

# punctuation
install.packages('redr'), correct
instll_packages('readr'), not correct


# R basics

variable <- function(argument)

#List of useful functions to create and explore data structure
* Basic mathematical functions for summary statistics
    + `sum()`
    + `mean()`
    + `min()`
    + `max()`
    + `sd()`
    + `median()`
    + `summary()`
    
* Functions to create & explore data structure 
    + `c()` - concatenate function
    + `:` - colon operator for sequence generation
    + `seq()` - sequence generation
    + `rep()` - replicate elements of vectors and lists
    + `View()` - invoke a data viewer
    + `length()` - length of an object
    + `class()` - object class
    + `head()` - return the first 6 rows of an object
    + `tail()` - return the last last 6 rows an object
    + `dim()` - returns the dimensions of an object 
    + `nrow()` - number of rows
    + `ncol()` - number of columns
    + `str()` - display the structure of each column of an object
    + `names()` - display the names of an object, typically the column names
    + `nlevels()` - categorial levels within an object
    + `cbind()` - combine objects by columns
    + `rbind()` - combine objects by rows
    + `rm()` - remove an object
    + `edit()` - edit an object
    + `t()` - transpose a matrix or dataframe
    + `is.data.frame()` - check if you data is a data frame
    + `as.data.frame()` - coerce your data to a data frame
    + `is.matrix()` - check if your data is a matrix
    + `as.matrix()` - coerce your data to a matrix




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages('readr') # if first time using readr
## hashtag is making a comment, it is not seen as code
# annotating your code is great practice! 

library(readr)
library(ggplot2)
```


#Exercises

```{r}
# A
bat_tongues <- read_csv("battongues.csv")

#C
bat_copy <- bat_tongues

names(bat_copy)

# let's change them
names(bat_copy) <- c("spp", "palate", "tongue")
names(bat_copy)

# mean of tongue
mean(bat_copy$tongue)

# other stats
summary(bat_copy$palate)

#### Caffeiene 

caffeine <- read_csv("caffeine.csv")
names(caffeine) <- c("brand", "caffeine_mg")


# basic stats
var(caffeine$caffeine_mg)
mean(caffeine$caffeine_mg)
max(caffeine$caffeine_mg)
min(caffeine$caffeine_mg)
sd(caffeine$caffeine_mg)
median(caffeine$caffeine_mg)
range(caffeine$caffeine_mg)
quantile(caffeine$caffeine_mg)

# or just run summary()
summary(caffeine$caffeine_mg)

# how to analyze categorical data
table(caffeine)
table(caffeine$brand)

# plot it!
ggplot(data = caffeine, aes(x = brand, y = caffeine_mg)) +
  geom_histogram(stat = "identity") +
  theme_minimal()+
  theme(axis.text.x.bottom =  element_text(size = 8, angle = 90))
  
```

#Homework

**Question 1.**
```{r}
# load in required data
mammal <- read_csv("mammals.csv")

# let's look at file
#View(mammal)

names(mammal) <- c("name", "body", "brain")


# A. Which mammal has the largest brain size:body ratio?

# make new variable column
mammal$ratio <- mammal$brain/mammal$body

# find max ratio
# dataframe[which.function(sf$variable),]

mammal[which.max(mammal$ratio),] # don't forget the comma post ()!!!

# B. Which mammal has the smallest ratio?
mammal[which.min(mammal$ratio),] # don't forget the comma post ()!!!

# C. What is the mean brain size:body ratio? Are humans above or below the mean?
mean(mammal$ratio)

# find human ratio
# Option 1 (prefered)
mammal[mammal$name == "Human",] # make sure the spelling is the exact same!

# Option 2
mammal[mammal$name == "Human", "ratio"] # if you have a lot of variable columns

# Option 3 (less prefered)
# make new dataframe for humans and analyze that
human <- mammal[which(mammal$name == "Human"),]
mean(human$ratio)
```
Answers
A. The ground squirrel has the largest brain size:body ratio of 40 g/kg.
B. The African elephant has the smallest brain size:body ratio of 0.858431 g/kg.
C. The mean brain size:body ratio is 9.516999 g/kg.  This is less than the human ratio of 21.29032.
D. Explain your findings about the scientist hypothesis in complete sentencees.

**Question 2**
```{r}
# load data

birds <- read_csv("desert_birds.csv")
#View(birds)

# A. How many desert bird species are rare (<= 10 counts)

# make subset
rare <- subset(birds, Count <= 10)
#find how many rows there are in subset
# Option 1. Look at length of vector
length(rare$Count)
dim(rare)
nrow(rare)

# Option 2. Look at dataset summary
print(rare)
summary(rare)

# This is different than finding the total sum of birds within these rare species
sum(rare$Count)
```
Answer. There are 16 species of desert birds that are considered rare. 

**Question 3**
```{r}

# load data
endangered <- read_csv("endangered.csv")

#View(endangered)

# A. What proportion of endangered species are vertebrates?

# Option 1

Verts <- subset(endangered, vertebrate == "yes")
sum(Verts$no.species)/ sum(endangered$no.species) # 0.26

# Option 2

verts <- sum(endangered$no.species[endangered$vertebrate == "yes"])
prop <- verts/sum(endangered$no.species)
print(prop) # 0.27

# Option 3
# make new dataframe

vertebrate <- subset(endangered, vertebrate == "yes")

# find sum of vertebrates
vertebrate_count <- sum(vertebrate$no.species)
print(vertebrate_count) # 399

# find sum of all endangered species (verts & inverts)
endangered_count <- sum(endangered$no.species)
print(endangered_count) # 1263

# find proportion of vertebrates
proportion <- vertebrate_count/endangered_count
print(proportion) # 0.27




```
Answer. 27% of endangered species are vertebrates.
