---
title: "Lab3_DiscussionNotes"
author: "sbsambado"
date: "1/21/2021"
output: html_document
---

Today's agenda

1. Class housekeeping notes
2. Work on exercises & homework in class
4. Ask questions


*Notes*

**If you get stuck on HW code**
+ 1. refer to course material
+ 2. Nectir
+ 3. Google
+ 4. Attend a lab section
+ 5. email a TA after 1 - 4

**Knitting problems**
+ make sure to have library()
+ #install.packages and #View
+ deceipher knitr error
+ read in data

**Don't hesitate to ask questions in section!**
+ often I recieve the same question from different students

**Check out these resources for graph help**
+ https://www.r-graph-gallery.com/
+ ggplot2-cheatsheet.pdf
  + hone in on graph types for numeric/categorical data

**Plotting**
+ base plot or ggplot - either is fine

+ ggplot formula

    ggplot(dataframe, aes(x = variable1, y = variable2)) +
    geom_structure()

**THE FORMULA IS Y ~ X**
+ variable of interest ~ explanatory variable
+ outcome ~ input


**Feedback on how you would like discussion section organized**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



```{r upload neccessary packages and data}
library(readr)
library(tidyverse)
library(ggplot2)

guppies <- read_csv("guppies.csv")
lynx <- read_csv("lynx_data.csv")
plant <- read_csv("plant_data.csv")

install.packages("ggplot2")
library(ggplot2)


```

# HOMEWORK

**Question 1**

```{r q1 hint}
# read in data
plant <- read.csv("plant_data.csv")

# look at data
#View(plant)
dim(plant)
summary(plant)

# answer question

# A. Describe the shape for different distances

par(mfrow = c(3,1)) # view 3 rows in 1 column

# HINT

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "0.0m"], xlab = "TITLE", ylab = "TITLE,  main = "TITLE)

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "3.0m"], xlab = "TITLE", ylab = "TITLE, main = "TITLE)

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "6.0m"], xlab = "TITLE", ylab = "TITLE, main = "TITLE")

par(mfrow = c(1,1)) # reset

# Make a boxplot to look at spread

#HINT
#boxplot(Y ~ X, DATAFRAME, xlab = "TITLE", ylab = "TITLE", main = "TITLE")

boxplot(plant$dand_rosette_diam_cm ~ plant$dist_from_edge_m, xlab = "Distance from edge", ylab = "diameter")

```

```{r q1 answer}
# A

par(mfrow = c(3,1))
# 0 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],xlab = "Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 0m")

# 3 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "3.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 3m")

# 6 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "6.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 6m")

par(mfrow = c(1,1))
# look at "tails", where are they pointing
# all look slightly skewed, with right skewed tails

# B
# Make a boxplot of dandelion rosette diameter
boxplot(dand_rosette_diam_cm ~ dist_from_edge_m, plant, xlab="Distance from edge",
		ylab="Dandelion rosette diameter")

# not confident true effect

par(mfrow = c(1,1))


```


**Question 2**

```{r q2 hint}
# A

# First, transform data (i.e. make 2 new columns in plant data frame)
# Then, make 3 histograms 
  # num_leaves
  # log_leaves
  # sq_leaves


# transform data


# HINT


# log transform
# DATAFRAME$LOG_VARIABLE = log(DATAFRAME$VARIABLE + 1)

# square transform 
# DATAFRAME$SQ_VARIABLE = DATAFRAME$VARIABLE^2

# look at distribution 

# HINT

par(mfrow = c(3,1))
#hist(DATAFRAME$VARIABLE,  xlab = "TITLE", ylab = "TITLE", main = "TITLE")
#hist(DATAFRAME$LOG_VARIABLE, xlab = "TITLE", ylab = "TITLE, main = "TITLE")
#hist(DATAFRAME$SQ_VARIABLE,  xlab = "TITLE", ylab = "TITLE, main = "TITLE")
```

```{r q2 answer}

plant$log_num_flowers = log(plant$num_flowers + 1)
plant$sq_num_flowers = plant$num_flowers^2


par(mfrow = c(3,1))
hist(plant$num_flowers, xlab="Number of flowers")
hist(plant$log_num_flowers, xlab="Log number of flowers")
hist(plant$sq_num_flowers, xlab="Number of flowers squared")
```

**Question 3**

```{r q3 hint}

plant <- read_csv("plant_data.csv")
# A. what trend do you notice in the data?

# HINT
# plot(DATAFRAME$VARIABLE1 ~ DATAFRAME$VARIABLE2, xlab = "TITLE", ylab = "TITLE, main = "TITLE")

# fit <- lm(DATAFRAME$VARIABLE1 ~ DATAFRAME$VARIABLE2, data = DATAFRAME)
# abline(fit, col = "CHOOSECOLOR")

```

```{r q3 answer}
# how you would code this in R
plot(plant$percent_cover ~ plant$dist_from_tree_m, xlab="Distance from tree", ylab="Percent cover")

fit <- lm(plant$percent_cover ~ plant$dist_from_tree_m, data=plant)
abline(fit,col="green")
```

