---
title: "Lab3_answers"
author: "sbsambado"
date: "1/16/2021"
output: html_document
---


# Answers for Lab 3

**Question 1** (3 points: 1 point for code, 1 point for correctly answering each answer)

+ a. All distances are slightly skewed, with tails pointing to the right (right-skewed). 3m and 0m look the most normal.  

+ b. The median rosette diameters are very similar at different distances from the
edge.  However, the dispersion or spread of the data increaes as the distance from the edge increases. Based on this, I would not feel confident concluding that there
is a significant effect of edge on median rosette diameter, but I would feel confident concluding that there is a significant effect of edge on the variance in rosette diameter.

**Question 2** (4 points: 1 point for code, 1 code for correctly answering each question)

+ a. Squaring the data only scaled the data, but did nothing about changing the distribution ("increasing the right skew"). Log-transforming the data changed the distribution ("decreased the right skew of the data"). 

+ b. The log-transformed data looks more normal.

+ c. The variable `num_flowers` can be 0 because it is a count. However the log of 0 is not a number; it is negative infinity.  Therefore, we need to add 1 to all values in the column to scale the data.  

Squaring zero still gives us zero. 

Technically, we would not need to add 1 if we log-transformed `dand_rosette_diam_cm` because measures are continuous variables that are always greater than 0. However, in our data set there actually are a few `dand_rosette_diam_cm` measures that are 0 due to mis-recording.


**Question 3** (3 points: 1 point for code and 1 point correctly answering each question)

+ a. As we increase the distance from the nearest tree we see an increase in the percent cover. However, the linear trend line is not a good representation of this trend. The trend looks distinctly non-linear and there is substantial variation in the data.
Just because R fitted a linear trend line doesn't mean that a line is the best fit. Even looking just looking at the slope of the line, we can tell it's a small relationship. Thus indicating that there is a weak to possibly no linear relationship.

+ b. The explanatory variable is `dist_from_tree_m` and the response variable is `percent_cover`.




*Notes*

+ ALL HOMEWORK SHOULD BE SUBMITTED AS A KNITTED HOMEWORK FILE (.PDF or .HTML)

  + -1.0 point for not PDF or HTML

+ - point for 1 day late
+ Take the time to nail in homework format since it's still early in course 
  + -.25 point if code is not annotated
  + -.25 point if answers are not in complete sentences
  + -.25 point if code does not seem adequate to answer question
  
+ no credit for corrupted/non-working file


# Some potential code to answer questions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(knitr)
library(ggplot2)
```

## Question 1 

```{r, fig.width = 5, fig.height = 6}

plant <- read.csv("plant_data.csv")

# Make a histogram of dandelion rosette diameter - using basic R plotting code
par(mfrow=c(3,1))
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 0m")
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "3.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 3m")
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "6.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 6m")
```


```{r}
# Make a boxplot of dandelion rosette diameter
boxplot(dand_rosette_diam_cm ~ dist_from_edge_m, plant, xlab="Distance from edge",
		ylab="Dandelion rosette diameter")
```

## Question 2 

```{r}
plant$log_num_flowers = log(plant$num_flowers + 1)
plant$num_flowers_sq = plant$num_flowers^2

```

Plotting the 3 histograms 

```{r, fig.width = 5, fig.height = 6}
# using using basic R plotting code
par(mfrow=c(3,1))
hist(plant$num_flowers, xlab="Number of flowers", breaks=20)
hist(plant$log_num_flowers, xlab="Log number of flowers", breaks=20)
hist(plant$num_flowers_sq, xlab="Number of flowers squared", breaks=20)

```
  

## Question 3 

```{r}
# how you would code this in R
plot(plant$percent_cover ~ plant$dist_from_tree_m, xlab="Distance from tree", ylab="Percent cover")

fit <- lm(plant$percent_cover ~ plant$dist_from_tree_m, data=plant)
abline(fit,col="green")
```


