---
title: "DiscussionNotes"
author: "sbsambado"
date: "4/13/2021"
output: html_document
---
Today's agenda

1. Where are we?
2. Class housekeeping notes
3. Work on exercises & homework in class
4. Ask questions

*Where are we?*

Week 1 -- hopefully everyone has R and Rstudio running on their computer
Week 2 -- Rmarkdown & data analysis
Week 3 -- Data visualization

# The Big Picture 

## 1. Data visualization is meant to convey a story. 
There are many ways to tell a story, but there are some general guidelines on how to build that story

  + What is your outcome of interest - your **dependent** variable 'y'? 
      + That outcome should be placed on the y-axis
  
  + What is influencing your outcome of interest - your **independent** variable 'x'? 
    + That should be placed on the x-axis
  
## 2. Your type of data determines how you shape your data 
  Type
        +  Numeric: numbers
        
        +  Categorical: words
        
  Scale
        +  Continuous: numbers can take on **any** value
        
        +  Discrete: numbers or words can only be **certain** values

## 3. Plotting basics
    + If you're only plotting 1 numeric variable --> histogram 
    
    + If y and x are **both** numeric --> scatterplot
    
    + If y is numeric, and x is categorical --> boxplot 

# 3. Data visualization is the **first** AND **last** step of statistical analysis

  +  Be prepared to make a lot of changes along the way & have fun with it!
  
  + Figures will never be 100% perfect in the creator's eyes, but if it tells your story in a **clear and effective** way, then that is a job *well done*.

Okay, end philosophical ramble. Let's get started. 


*Notes*

**If you get stuck on HW code**
+ 1. Refer to course material
+ 2. Nectir
+ 3. Google
+ 4. Attend a lab section
+ 5. email a TA after 1 - 4

**Knitting problems**
+ make sure to have library()
```{r}
# error: can't find read_csv function

library(readr)

# good coding practice, in your first chunk of your .Rmd file
# upload your libraries
# then upload your data 
```

+ #install.packages and #View
```{r}
 # install.packages()
 # View()

# errors: quartz, rtools, CRAN 
```

+ decipher knitr error
+ read in data

**Don't hesitate to ask questions in section!**
+ often I receive the same question from different students

**Check out these resources for graph help**
+ https://www.r-graph-gallery.com/
+ ggplot2-cheatsheet.pdf
  + hone in on graph types for numeric/categorical data

**Plotting**
+ base plot or ggplot - either is fine
```{r}

# new package alert
install.packages("ggplot2") # install if this is your first time using it
library(ggplot2)

# don't need to install anything for base plot

# let's read in some practice data to illustrate
plant <- read_csv("plant_data.csv")

### arguments
# argument 1

# HINT: ggplot(dataframe) 
ggplot(plant)

# argument 2
# HINT: ggplot(dateframe, aes(x = VARIABLEX, y = VARIABLEY)) # data you want to plot
ggplot(plant, aes(x = dist_from_edge_m, y =num_flowers ))

# argument 3
# HINT: ggplot(dateframe, aes(x = VARIABLEX, y = VARIABLEY)) +
  # geom_SHAPE()
ggplot(plant, aes(x = dist_from_edge_m, y =num_flowers )) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "distance") #shape of your data


```


+ ggplot formula

    ggplot(dataframe, aes(x = variable1, y = variable2)) +
    geom_shape()

**THE FORMULA IS Y ~ X**
+ variable of interest ~ explanatory variable
+ outcome ~ input


**Feedback on how you would like discussion section organized**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



```{r upload neccessary packages and data}

# packages for this week
library(readr) # allows you to read in packages
## new pacakge alert!
# install.packages("package")
library(tidyverse) # tidys your data
library(ggplot2) # plots your data

# datasets for this week
guppies <- read_csv("guppies.csv")
lynx <- read_csv("lynx_data.csv")
plant <- read_csv("plant_data.csv")
birds <- read_csv("desert_birds.csv")

```

Explore your data
```{r}
# structure of dataset


# details of dataset



# summary of dataset


```

# HOMEWORK

**Question 1**

```{r q1 hint}
# read in data
plant <- read.csv("plant_data.csv")

# look at data

dim(plant) # 259 observations/rows , 10 columns/variables

# View(plant)

str(plant)

```


```{r}
# go to for looking at dataset before analyzing or plotting 
# View(plant)
dim(plant) # 259 observations/rows and 10 columns/variables 
summary(plant)
str(plant)

# basic format for histogram 
# hist(DATASET$VARIABLEOFINTEREST)
hist(plant$dand_rosette_diam_cm)
hist(plant$dand_rosette_diam_cm) # diameter for all rosettes

```

```{r}

# but what is the diameter across the three levels of distance from edge?
# we're going to make 3 histograms, each representing either 0m, 3m, 6m
# so how do we select for each distance?

# square brackets [ ] select
# regular ( ) for normal functions and arugments

# when you're selecting for things, use == not = 

limited <- plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"]
#str(limited)



# (row, columns)
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],)

hist(plant$dand_rosette_diam_cm) # full column
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],) # specified colum
#hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.3m"],) # specified colum
#mhist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.6m"],) # specified colum

#hist(plant$dand_rosette_diam_cm)

# answer question

# currently this 3 window outputs
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],)
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "3.0m"],)
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "6.0m"],)

# make 1 window output
par(mfrow = c(3,1)) # changes my parameters to 3 rows and 1 column output
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],)
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "3.0m"],)
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "6.0m"],)
# make sure to re-run par function 
par(mfrow = c(1,1))


```

```{r}
# question 1b
# format: boxplot(dataset$variableY)
boxplot(plant$dand_rosette_diam_cm)
boxplot(plant$dand_rosette_diam_cm) # boxplot for all rosette diameters

# format: boxplot(dataset$variableY ~ dataset$variableX)
# Y ~ X
# outcome of interest ~ explanatory input
# dependent ~ independent variable
# Y ~ X
boxplot(plant$dand_rosette_diam_cm ~ plant$dist_from_edge_m)

# HINT: boxplot(Y ~ X)
boxplot(plant$dand_rosette_diam_cm ~ plant$dist_from_edge_m)

# unique() helps you understand the different levels/groupings within one variable 
# unique(plant$dist_from_edge_m)

# A. Describe the shape for different distances

par(mfrow = c(3,1)) # view 3 rows in 1 column

# HINT

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "0.0m"], xlab = "TITLE", ylab = "TITLE,  main = "TITLE)

hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],main = "Distance from 0m", xlab = "distance from edge", ylab = "diameter")

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "3.0m"], xlab = "TITLE", ylab = "TITLE, main = "TITLE)

#hist(DATAFRAME$VARIABLE1[DATAFRAME$VARIABLE2 == "6.0m"], xlab = "TITLE", ylab = "TITLE, main = "TITLE")

par(mfrow = c(1,1)) # reset

# Make a boxplot to look at spread

#HINT
#boxplot(Y ~ X, DATAFRAME, xlab = "TITLE", ylab = "TITLE", main = "TITLE")

boxplot(plant$dand_rosette_diam_cm ~ plant$dist_from_edge_m, xlab = "Distance from edge", ylab = "diameter")

```

```{r q1 answer}
# A

par(mfrow = c(3,1))
# 0 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "0.0m"],xlab = "Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 0m")

# 3 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "3.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 3m")

# 6 meters
hist(plant$dand_rosette_diam_cm[plant$dist_from_edge_m == "6.0m"], xlab="Dandelion diameter (cm)", ylab="Density", main = "Histogram of dandelion rosette diameter - 6m")

par(mfrow = c(1,1))
# look at "tails", where are they pointing
# all look slightly skewed, with right skewed tails

# B
# Make a boxplot of dandelion rosette diameter
boxplot(dand_rosette_diam_cm ~ dist_from_edge_m, plant, xlab="Distance from edge",
		ylab="Dandelion rosette diameter")

# not confident true effect

par(mfrow = c(1,1))


```


**Question 2**


the jist = transform your data

transforming data is sometimes necessary **especially in ecology**
- totally fine, you are scaling all of your data in the same way

3 graphs
1) regular data, just plot it as is
2) log transformed 
3) square transformation


Then you stack graphs 1 - 3 similar to question 1
Describe which graph is the most normal, aka the best bell curve shape 

**Error for question 2c**: square root == squared, but just choose the graph that has the best bell curve
```{r}
# to add a new column of log transformed data 
plant$log_num_flowers <- log(plant$num_flowers + 1)

# plant$sq_num_flowers
```

```{r q2 hint}
# A

# First, transform data (i.e. make 2 new columns in plant data frame)
# Then, make 3 histograms 
  # num_leaves
  # log_leaves
  # sq_leaves


# transform data


# HINT


# log transform
# DATAFRAME$LOG_VARIABLE = log(DATAFRAME$VARIABLE + 1)

# square transform 
# DATAFRAME$SQ_VARIABLE = DATAFRAME$VARIABLE^2

# look at distribution 

# HINT

par(mfrow = c(3,1))
#hist(DATAFRAME$VARIABLE,  xlab = "TITLE", ylab = "TITLE", main = "TITLE")
#hist(DATAFRAME$LOG_VARIABLE, xlab = "TITLE", ylab = "TITLE, main = "TITLE")
#hist(DATAFRAME$SQ_VARIABLE,  xlab = "TITLE", ylab = "TITLE, main = "TITLE")
```

```{r q2 answer}

plant$log_num_flowers = log(plant$num_flowers + 1)
plant$sq_num_flowers = plant$num_flowers^2


par(mfrow = c(3,1))
hist(plant$num_flowers, xlab="Number of flowers")
hist(plant$log_num_flowers, xlab="Log number of flowers")
hist(plant$sq_num_flowers, xlab="Number of flowers squared")
```

**Question 3**

```{r q3 hint}

plant <- read_csv("plant_data.csv")
# A. what trend do you notice in the data?

# HINT
# plot(DATAFRAME$VARIABLE1 ~ DATAFRAME$VARIABLE2, xlab = "TITLE", ylab = "TITLE, main = "TITLE")

# fit <- lm(DATAFRAME$VARIABLE1 ~ DATAFRAME$VARIABLE2, data = DATAFRAME)
# abline(fit, col = "CHOOSECOLOR")

```

```{r q3 answer}
# how you would code this in R

# 3 layers
# first you need a plot of real data (i.e. your rows in dataset are open circles)
plot(plant$percent_cover ~ plant$dist_from_tree_m, xlab="Distance from tree", ylab="Percent cover")
# simulated data (line) to support real world data (open circles)
fit <- lm(plant$percent_cover ~ plant$dist_from_tree_m, data=plant)

# add simulated fitted line
abline(fit,col="green")
```

