---
title: "Lab4_discussionNotes"
author: "sbsambado"
date: "1/28/2021"
output: html_document
---


Today's Agenda

1. What you need for this lab

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(ggplot2) # ifyou want to make fancy plots

library(car) # new package alert!

# data needed for this lab
titanic <- read_csv("titanic.csv")
dexterity <- read_csv("dexterity.csv")
bumpus <- read_csv("bumpus.csv")
```

2. Contextualize where we are this week
    Lab 2: import and wrangle data
    Lab 3: visualize data
    Lab 4: determine normality of data
    Lab 5 - 8: running the appropriate statistical tests based on your data normality
    
3. Quick Notes about Lab 4
    + it is A LOT of information 
        + but this information will help you with 30% of your final project
    + sometimes it is a judgment call based on hist/box, qqPlot, CLT
        + just make your an

4. Info that will make this lab assignment easier (this is a notoriously tricky lab)     

### shapiro wilks comparing NORMALITY
### t-test comparing MEANS
### levene's and welch's comparing VARIANCE

This lab is split into 2 sections
~ make sure you null/alternative hypotheses match with what tests you are running!! ~

**1. Normality**
  + shapiro wilks: testing normality 
  
**2. Comparing means**
  + t-test : testing for means
  + levene's / welch : testing for variance

# Shapiro-Wilks: Testing for normality 

*Steps to complete homework*
1. state your null and alternative hypothesis
2. visualize the distribution with a histogram/boxplot, qqPlot
3. run a shapiro-wilk test (test normality)
4. run the appropriate t-test (parametric or not parametric)


## WRITE THIS DOWN SOMEWHERE

*The gist of Shapiro-Wilks Test*
null hypothesis: data are NORMALLY distributed
alternative hypothesis: data are NOT NORMALLY distributed

p < 0.05 : reject null --> data are NOT normally distributed
p > 0.05 : fail to reject null --> data are NORMALLY distributed


# How to interpret shapiro test results

```{r}
# create data from a norm sample
norm_samp = rnorm(1000, mean=21 , sd=3) #normal distribution 

# run a shapiro.test
# HINT: shapiro.test(VECTOR)
shapiro.test(norm_samp)

shapiro.test(norm_samp)


```

  1. state your p-value: p-value = 0.348
  2. compare your p-value to significant p-value value: 0.348 > 0.05
  3. look at your null hypothesis: data are normally distributed
  4. contextualize: My p-value of the shapiro-wlk test is 0.348, which is above a p-value of 0.05, so I fail to reject my null and I can be resaonably confident that my data are normal

*Does your shaprio-wilk value align with the plots you made regarding data distribution?*
+ this is a judgment call sometimes, when in doubt look at CLT (50+ observations)

run a qqPlot for normality
```{r}

# other tools for visualizing normality 
# histogram
# boxplots

# another tool for visualizing normality
# qqPlot

qqPlot(norm_samp) 
```

# T-test: Comparing Means

paired = take the difference (requires subsetting)
  + can't split the groups
    + left/right hand 1 body, leaf 1/leaf2 1 branch

two - sample = splits 1 variable into 2 groups with code
  + can split the groups
    + high school/college students, basketball/baseball players
