---
title: "lab4_DiscussionNotes_S21"
author: "sbsambado"
date: "4/20/2021"
output: html_document
---

Today's Agenda

1. *What you need for this lab*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# go to packages
library(readr) # read.csv function
library(tidyverse) # make tidy data
library(ggplot2) # if you want to make fancy plots

# new package alert!
# install.packages("car")
# install.packages("car")
library(car) # visualize with qqPlot

# data needed for this lab
titanic <- read_csv("titanic.csv")
dexterity <- read_csv("dexterity.csv")
bumpus <- read_csv("bumpus.csv")
```

2. *Where are we?*

Week 1 -- hopefully everyone has R and Rstudio running on their computer
Week 2 -- Rmarkdown & data analysis
Week 3 -- Data visualization
Week 4 -- Determine normality & compare means


3. *Quick Notes about Lab 4*
    + it is A LOT of information 
        + but this information will help you with 30% of your final project
    + sometimes it is a judgment call based on hist/box, qqPlot, CLT
        + just make your answer complete and logical

4. *Info that will make this lab assignment easier (this is a notoriously tricky lab)*     

### shapiro wilks comparing NORMALITY
### t-test comparing MEANS
### levene's and welch's comparing VARIANCE

This lab is split into 2 sections
~ make sure you null/alternative hypotheses match with what tests you are running!! ~

**1. Normality**
  + shapiro wilks: testing normality 
  
**2. Comparing means**
  + t-test : testing for means
  + levene's / welch : testing for variance

# Shapiro-Wilks: Testing for normality 

*Steps to complete homework*
1. state your null and alternative hypothesis
2. visualize the distribution with a histogram/boxplot, qqPlot
3. run a shapiro-wilk test (test normality)
4. run the appropriate t-test (parametric or not parametric)


## WRITE THIS DOWN SOMEWHERE

*The gist of Shapiro-Wilks Test*
null hypothesis: data are NORMALLY distributed
alternative hypothesis: data are NOT NORMALLY distributed
```{r}

# HINT: shapiro.test(DATASET$VARIABLE)
shapiro.test(titanic$Age)
```

p < 0.05 : reject null --> data are NOT normally distributed
p > 0.05 : fail to reject null --> data are NORMALLY distributed


# How to interpret shapiro test results

```{r}
# create data from a norm sample
norm_samp = rnorm(1000, mean=21 , sd=3) #normal distribution 

# run a shapiro.test
# HINT: shapiro.test(VECTOR)
shapiro.test(norm_samp)

shapiro.test(norm_samp)

qqPlot(norm_samp)
shapiro.test(norm_samp)
```

  1. state your p-value: p-value = 0.472
  2. compare your p-value to significant p-value value: 0.472 > 0.05
  3. look at your null hypothesis: data are normally distributed
  4. contextualize: My p-value of the shapiro-wlk test is 0.348, which is above a p-value of 0.05, so I fail to reject my null and I can be resaonably confident that my data are normal

*Does your shaprio-wilk value align with the plots you made regarding data distribution?*


testing for normality, use 4 tools
- visualize with boxplot or historgram
- visualize with a qqPlot
- run a shapiro wilk test
- see if the CLT applies (50+ observations or rows = data are normal)


+ this is a judgment call sometimes, when in doubt look at CLT (50+ observations)

run a qqPlot for normality
```{r}

# other tools for visualizing normality 
# histogram
# boxplots

# another tool for visualizing normality
# qqPlot is in the car package 

library(car)


qqPlot() # right


qqplot() # wrong
qqp() # wrong

# HINT: qqPlot(DATASET$VARIABLE)
qqPlot(titanic$Age)

#View(titanic)
```

# T-test: Comparing Means

*paired* = take the difference (requires subsetting)
  + can't split the groups
    + left/right hand 1 body, leaf 1/leaf2 1 branch

*two - sample* = splits 1 variable into 2 groups with code
  + can split the groups
    + high school/college students, basketball/baseball players
```{r}
# Option 1 for making two variables
# function subset it comes from the package tidyverse 
View(dexterity)

# function subset, you need the package tidyverse

library(tidyverse)

# HINT : variable name <- subset(DATASET, COLUMN == "TYPE")
dominant <- subset(dexterity, hand=="dominant")
weak  <- subset(dexterity, hand=="weak")


# Option 1: paired t-test
t.test(dominant, weak, paired = TRUE)

# make a graph for both dominant and weak 


# Option 2: one sample t-test
diff <- dominant$time - weak$time
t.test(diff, mu = 0)
# visualize and test normality on the diff!

hist(diff)
shapiro.test(diff)

boxplot(dexterity$time ~ dexterity$hand)

# paired t-test = weak and dominant (visualize boxplot or histogram)
t.test(weak, dominant, paired = TRUE)

# one sample t-test = diff (visualize boxplot or histogram)
t.test(diff, mu = 0)
```

```{r}
leveneTest(total_length_mm ~ survival, data = bumpus, center = "median")

# ?leveneTest
# str(bumpus)

t.test(total_length_mm ~ survival, data = bumpus, var.equal = TRUE)
```

visualizing with boxplot/histogram/qqPlot

- mention outliers 
- normal distribution (bell curve or not)
- varaince
```{r}

with(dexterity, qqPlot(time[hand == "dominant"]))
with(dexterity, qqPlot(time[hand == "weak"]))


with(bumpus, qqPlot(total_length_mm[survival == "no"]))
with(bumpus, qqPlot(total_length_mm[survival == "yes"]))

# HINT: boxplot(DATASET$Y ~ DATASET$X)
boxplot(bumpus$total_length_mm ~ bumpus$survival)

```

