---
title: "lab5_discussionnotes"
author: "sbsambado"
date: "2/4/2021"
output: html_document
---

Today's Agenda

1. What you need for this lab

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries needed for this lab
library(readr)
library(tidyverse)
library(ggplot2) # ifyou want to make fancy plots
library(car) 

mammals <- read_csv("mammals_hearing.csv")
parasitoids <- read_csv("parasitoids.csv")
crickets <- read_csv("starving_cricket.csv")
plant <-read_csv("plant_data.csv")


```

2. Contextualize where we are this week
    Lab 2: import and wrangle data
    Lab 3: visualize data
    Lab 4: determine normality of data
    Lab 5: what to do with non-normal data... moving into regression models!
    
3. Quick Notes about Lab 5
    + You know most of this stuff already.
    + residuals are *new* 
        ~ amount of variability between observed and predicted Y value
            aka "how wrong is your prediction"
      res = 0 -> no difference between your observed & predicted Y values (highly unlikely)
      res = 1+ -> there is some difference between observed & predicted Y values (more likely)
    
    + Use your lab manuals to follow the structure of how we want you to answer question
    + Be *very* clear with your language 
  
4. Info that will make this lab assignment easier (this is a notoriously tricky lab)     
  + parametric for normal data
  + non-parametric for not normal data
  
    **shapiro wilks** comparing NORMALITY
    **t-test** comparing MEANS
    **levene's and welch's** comparing VARIANCE
    **Mann-Whitney U-test/Wilcoxn** comparing MEDIANS (non-parametric) << this is new
    
*The gist of these tests *
null hypothesis: data are X 
alternative hypothesis: data are NOT X

p < 0.05 : reject null --> data are NOT X
p > 0.05 : fail to reject null --> data are X


MAKE SURE IF YOU TRANSFORM YOUR DATA YOU RE-RUN THROUGH ALL THE NORMALITY CHECKS
+ hist(data) -> not normal
  qqPlot(data) -> not normal
  shapiro(data) -> not normal
  
  
+ hist(log(data)) -> normal
  qqPlot(log(data)) -> normal
  shapiro(log(data)) -> normal
      ----> then run all your appropriate statistical tests!
 
```{r}

unique(plant$dist_from_edge_m)
unique(plant$num_leaves_in_rosette)

??subset()


boxplot(num_leaves_in_rosette ~ dist_from_edge_m, data = sub_plant)

qqPlot(num_leaves_in_rosette ~ dist_from_edge_m, data = sub_plant)


```

