---
title: "lab6_disccusion_notes"
author: "sbsambado"
date: "2/11/2021"
output: html_document
---

Today's Agenda

1. FINAL PROJECT
+ go over the instructions
+ start looking for datasets **NOW**
+ look at link `How to start looking at data`


2. What you need for this lab
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages needed for lab 6
library(readr)
library(tidyverse)
library(car)
library(ggplot2)
library(multcomp) # new package alert! Make sure to install.packages("")
library(MASS) # new package alert! Make sure to install.packages("")

# datasets needed for lab 6
hand_wash <- read_csv("hand_washing.csv")
malaria <- read_csv("malaria_vs_maize.csv")
eggs <- read_csv("cuckooeggs.csv")

# you'll need `power_anova.R` open in another script



```

3. Contextualize where we are this week
    Lab 2: import and wrangle data
    *T-test land* -> Part 1 of final project
    Lab 3: visualize data
    Lab 4: determine normality of data
    Lab 5: what to do with non-normal data... moving into regression models!
    *Regression and correlation land* -> Part 2 of final project
    Lab 6: Regression models are here and residuals are here to stay
    
4. Quick Notes about Lab 6
  + analyzing variance within and across groups
  
  + **factor** = variable of interest (column name)
      + **level/treatment** = groups we are comparing
      
  + MAKE SURE YOUR VARIABLE IS A FACTOR
```{r}
hand_wash < read_csv("hand_washing.csv")
str(hand_wash)

# transform washing_method into a factor

# DATASET$FACTOR <- as.factor(DATASET$FACTOR)
hand_wash$washing_method <- as.factor(hand_wash$washing_method)
str(hand_wash) # washing_method = 'chr', won't work for the mcp()
```

  + HW question 3 copy script `power_anova.R`
    + only change values in top part of script
    

  + How to read an ANOVA residual plot
```{r}
# HINT: aov(y ~ x, data = dataset)
wash1 <- aov(bac_colonies~washing_method, data=hand_wash) #run the ANOVA
par(mfrow=c(2,2)) # places 2 plots on 1 output view
plot(wash1) #look at the diagnostic plots; focus on the first two plots (fitted vs residuals & QQ plot)

unique(hand_wash$washing_method)
```
+ Different wrappers
```{r}
### ONE WAY ANOVA ### 

# run model with aov()
# get summary of model with summary()

# run model with lm()
# get ANOVA summary of model with aov()

```


You test residuals to get the whole picture
```{r}
par(mfrow = c(3,1))
with(plant,hist(num_leaves_in_rosette[dist_from_edge_m=="0.0m"], breaks = 20, col = "lightblue"))
with(plant,hist(num_leaves_in_rosette[dist_from_edge_m=="6.0m"], breaks = 20, col = "darkblue"))
with(plant,hist(num_leaves_in_rosette, breaks = 20, col = "blue"))


?hist()
```

