---
title: "lab8 discussion notes"
author: "sbsambado"
date: "2/25/2021"
output: html_document
---

Today's Agenda

### 1. FINAL PROJECT

- Mini-Assessment # 2 is due tomorrow at 12 PM (NOON)
      4 tools to check for normality
      - boxplot/histogram
      - shapiro wilk
      - qqPlot
      - CLT
        >> then you make the final call!
        
- My sections will now be focused on final project help

**Your questions boils down to this**
(Y) numeric/normally distributed ~  (X) numeric + (X) categorical + (X) categorical

- Use the tools and language we have gone over in lecture and lab
- This is a professional report

If you are stuck on where to begin

- look at structure of data
```{r}
dim(dataset) # observations, variables
str(dataset) # type of data (numeric vs categorical)
dim(dataset) # observations, variables
summary(dataset) # summaries of numeric variable s(mean/min/max, etc)
```

- find important variables

- start plotting (exploratory data analysis = EDA)
```{r}

y + x1 + x2 + x3


boxplot or scatter of y ~ x1
boxplot or scatter of y ~ x2
boxplot or scatter of y ~ x3

# boxplot
#histogram
# ggplots
```

- find relationships that are similar to what we've done in class

### 2. What you need for this lab
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages needed for lab 8
# If you haven't used a package before you will need to install it with install.packages("package") 
library(readr)
library(tidyverse)
library(ggplot2)
library(car) #for the qqPlot function
library(psych) #for pairs.panels()

library(knitr) # we will use this package to make beautiful tables
library(nlme) #for fitting mixed effects models (linear models with random effects) ~ new pacakge alert! ~

# datasets needed for lab 8
plant<- read_csv("plant_data.csv") # obviously
airqual <- airquality # this is a built in data set in R, type `data()` into your console to see all the datasets there are!


```

### 3. Contextualize where we are this week
    Lab 2: import and wrangle data
    Lab 3: visualize data
    *Comparing means/medians/variance land* -> Part 1 of final project (t-test or ANOVA)
    Lab 4: determine normality of data
    Lab 5: what to do with non-normal data
    Lab 6: how to assess variance of categorical data (check normality on residuals)
    *Regression and correlation land* -> Part 2 of final project
    Lab 7: regression and correlation
    Lab 8: multiple regression and model selection **THIS IS THE MOST USEFUL LAB**
    
###  4. Quick Notes about Lab 8

  If we want to predict things
    *y ~ x* doesn't cut it most times
    *y ~ x + x + x* is sometimes better
  
  **Remember rules of parsimony, keep it simple.**
  
  child's height ~ mom height + nutrition score (two Xs)
  child's height ~ mom height + length of mom's jeans (two Xs)
  
  **Be mindful of the difference between lm() and lme()**
  - you may have difficulty assembling the table if you leave in your lme() model
      
      + lm can take on the forms
              *linear regression*: y ~x_1
              *multiple regression*: y ~ x_1 + x_2 + x_3
              *multiple regression with an interaction*: y ~ x_1 * x_2
              
      + lme is an extension of lm, but now with mixed effects
              *linear regression*: y ~ fixed = x_1, *random = (1|x_2)*
              *multiple regression*: y ~ x_1 + x_2 + x_3, r*andom = (1|x_4)*
              *multiple regression with an interaction*: y ~ x_1 * x_2 , *random = (1|x_3)*
      
check normality of residuals
>> residuals = observed y/outcome - predicted y/outcome

approach data
perform normality checks of your residual (untransformed y) (boxplot, qqPlot, shapiro)
if not normal
transform y (log, sqrt, arc)
re-run your normality check (transformed y )



Let's go over pairs.panel one more time
```{r message=FALSE}
# clean up your data
plant <- read_csv("plant_data.csv")
str(plant) # 259 observations, 10 variables
plant$tran_number <- as.factor(plant$tran_number)
plant$day_of_week <- trimws(plant$day_of_week, which = "both") #this gets rid of the spaces around the words
plant$day_of_week <- as.factor(plant$day_of_week)
plant <- na.omit(plant)

# run pairs.panel


pairs.panels(plant, lm = TRUE, cor = T, density = T)
pairs.panels(plant, lm = TRUE, cor = T) #note - num_flowers and num_dand_flowers are highly correlated

plant_sub <- subset(plant, select = -num_dand_flowers)

pairs.panels(plant_sub, lm = TRUE) #now there are no very strong correlations



```

    
Let's go over how to read model summary outputs one more time
```{r}
fit_2var <- lm(percent_cover ~ num_leaves_in_rosette + dist_from_edge_m, 
               data = plant_sub) 

summary(fit_2var)

qqPlot(fit_2var$residuals)

# p-value = 1.013e-08
# Intercept (B0) = 70.28
# F statistic = 14.67 
# Adjusted R square = 0.1596

# if you want to figure out how many levels you have within a factor
# HINT: unique(DATASET$FACTOR)

??HSD

```

```{r}
```

