---
title: "lab8 answers"
author: "sbsambado"
date: "2/21/2021"
output: html_document
---


# Answers for Lab 8

**Question 1**
1 point for background/context, -0.5 for uncited sources

**Question 2**
0.5 point for visualization, 0.5 for identification of predictors (SolarR, Wind, and Temp could all be predictors; Month could be justified as predictor or random effect; Day could be random effect but may not be important)

**Question 3**
0.5 point for checking that there is a linear relationship between predictors and response. 0.5 point for addressing outliers (points 117 and possibly 62; either run model with/without them or log transform Ozone) remember, not all predictors need to be normal as long as the model residuals are normal

**Question 4**
1 point for linear model with correctly checking residuals

**Question 5**
2 points (each model, with correctly checked residuals, is 1 point)

**Question 6**
1 point for nicely formatted table with df, AIC, BIC, and adj.r.sq

**Question 7**
1 point for justification of model choice; 1.5 point for prediction/assessment of fit

0.5 point for formatting/style/overall professional appearance of assignment

*Notes*

ALL HOMEWORK SHOULD BE SUBMITTED AS A KNITTED HOMEWORK FILE (.PDF or .HTML)

  + -2.0 points if HW not in right format 
  + -.25 point if code is not annotated
  + -.25 point if answers are not in complete sentences
  + -.25 point if code does not seem adequate to answer question



```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages needed for lab 8
# If you haven't used a package before you will need to install it with install.packages("package") 
library(readr)
library(tidyverse)
library(ggplot2)
library(car) #for the qqPlot function
library(psych) #for pairs.panels()
library(knitr) # we will use this package to make beautiful tables
library(nlme) #for fitting mixed effects models (linear models with random effects) ~ new pacakge alert! ~

# datasets needed for lab 8

airqual <- na.omit(airquality)

```

**Question 1**
Pretty much any background info is fine.

**Question 2: Visualize**
```{r}
#must use pairs plots
pairs.panels(airqual, lm = T) #discuss which variables have linear relationships

#optional: boxplot - remember this doesn't have to be normal
boxplot(airqual$Ozone)

```

**Question 3: Check assumptions**
```{r}
fit_solar <- lm(Ozone~Solar.R, data = airqual)
fit_wind <- lm(Ozone~Wind, data = airqual)
fit_temp <- lm(Ozone~Temp, data = airqual)
fit_all <- lm(Ozone~Temp+Wind+Solar.R, data = airqual)

res_S <- resid(fit_solar)
res_W <- resid(fit_wind)
res_T <- resid(fit_temp)
res_A <- resid(fit_all)

op <- par(mfrow = c(2,2))
qqPlot(res_S)
qqPlot(res_W)
qqPlot(res_T)
qqPlot(res_A)

par(op)

logfit_solar <- lm(log(Ozone)~Solar.R, data = airqual)
logfit_wind <- lm(log(Ozone)~Wind, data = airqual)
logfit_temp <- lm(log(Ozone)~Temp, data = airqual)
logfit_all <- lm(log(Ozone)~Temp+Wind+Solar.R, data = airqual)

logres_S <- resid(logfit_solar)
logres_W <- resid(logfit_wind)
logres_T <- resid(logfit_temp)
logres_A <- resid(logfit_all)

op <- par(mfrow = c(2,2))
qqPlot(logres_S)

par(op)

```

**Question 4,5**
Any models are fine
```{r}
#random effects example
re_model <- lme(log(Ozone) ~ Temp, random = ~1|Month, data = na.omit(airqual))
re_res <- resid(re_model)
qqPlot(re_res)

#interaction example
x_model <- lm(log(Ozone) ~ Temp*Wind, data = airqual)
x_res <- resid(x_model)
qqPlot(x_res)

#check models with and without outliers
sub_airq <- dplyr::filter(airqual, Ozone != 1) #better way to get rid of row 21?
x_model_SUB <- lm(log(Ozone) ~ Temp*Wind, data = sub_airq)
qqPlot(resid(x_model_SUB))
```

**Question 6**
```{r}
result <- AIC(logfit_solar,logfit_all, logfit_temp, logfit_wind, x_model) #this will create a dataframe whose rownames are the models, with columns for the df and AIC of each model # don't run re_model cause made with lme

#add other metrics to your table
models <- list(logfit_solar,logfit_all, logfit_temp, logfit_wind, x_model) #make sure you keep your models in the same order here as they were when you created your results table
result$BIC <- sapply(models, BIC) #add a column for BIC to the results

#result$AICc <- result$AIC+ (2*(result$df-2)^2+2*(result$df-2))/(153-1-(result$df-2))
#AICc doesn't make a difference bc sample size is big enough

model_summary <- lapply(models, summary) #look up ?lapply if you have not used this function before

#now we will use a for loop to easily extract the R^2 and adj R^2 value for each model from its summary, and store them in new columns in the results table

for(i in 1:(length(models)-1)){ #this creates a variable i that starts with the value i=1
  result$rsq[i] <- model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result$adj_rsq[i] <- model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again


kable(result, digits = 2, align = "c")
```

**Question 7**
Best model here is logfit_all based on AIC, BIC, R2 but it doesn't do that well predicting
```{r}
splitter <- sample(1:nrow(airqual), 10, replace = F) #pick 15 random rows from plant_sub to reserve as test data
airq_train <- airqual[-splitter,] #leave those rows out of the training data
airq_test <- airqual[splitter,] #use them to create a set of test data

final_fit <- lm(Ozone ~ Solar.R + Wind + Temp, data = airq_train) #fit the final model, using JUST your training set as the data argument

prediction <- predict(final_fit,airq_test) #use the fitted model to predict values for your test data

plot(airq_test$Ozone, pch=1) #plot the actual test data values
points(prediction, pch=20, col = "red") #plot the model predictions for those points
```

