---
title: "R tips & tricks (full)"
author: "sbsambado"
date: "12/23/2020"
output: html_document
---

##Tutorial Video: Making a R Project (.Rproj) for EEMB146

#Why? To find files & keep file paths simple.

**You'll hear about setwd()**
+ Which takes an absolute file path as an input then sets it as the current working directory of the R process.

+ You then use getwd() to find out what the current working directory is, and check that your working directory is correctly set.

```{r example of setwd}
setwd("/Users/jenny/cuddly_broccoli/verbose_funicular/foofy/data")
getwd()
df <- read.delim("raw_foofy_data.csv")
```

^^ this is fine... but,
+ What if you want to run other people's code?
+ What if you are working on multiple projects?

And most realisticly, you will run into 
>Error in file(file, "rt") : cannot open the connection

**There is a better way**

### R Projects are your friend for this course, other courses and any additional projects you may be working on

## Why? R Projects create easy file path referencing within RStudio

This .Rproj file can be created by going to File > New Project… in RStudio, 
+ which then becomes associated with the specified folder or directory. 
+ The mindset should then be that the directory (the whole folder and its sub-folders and contents) is stand-alone and portable, 
+ which in other words means that you shouldn’t be reading in data from or writing data to files outside the directory. 
+**Everything relating to that analysis or project should only happen within that directory**


**sources: <https://www.tidyverse.org/blog/2017/12/workflow-vs-script/>**



##Tutorial Video: How to upload data




First watch: Lab1, intro to R and Rstudio


First things first, let's get acquainted with the file types in RStudio
here is a good link if you're interested
https://martinctc.github.io/blog/rstudio-projects-and-working-directories-a-beginner's-guide/

Types of files you can work with and export
A) filename.R (R Script - don't use this)
+ like a word doc 
B) filename.Rmd
C) filename.html
D) filename.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
